Explore Local Memory Design for Building Hypervisor on Resource Disaggregated Architecture

### 2019-11-03 paper.tex
Add introduction and background.

% Figure of monolithic architecture and disaggregared architecture
\begin{figure*}[h!]
     \centering
     \captionsetup[subfigure]{position=b}
     \begin{subfigure}[b]{0.28\textwidth}
         \includegraphics[height=.2\textheight]{figures/monolithic.png}
         \caption{Monolithic architecture, hardware components are bounded within a machine.}
         \label{fig:monolithic architecture}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.68\textwidth}
         \includegraphics[height=.2\textheight]{figures/disaggregated.png}
         \caption{Disaggregated architecture, hardware components are separate: there are CPU-blade, memory-blade, storage-blade. We also assume there is local memory located with CPU}
         \label{fig:disaggregated architecture}
     \end{subfigure}
     \caption{Monolithic architecture and disaggregated architecture.}
     \label{fig:architecture}
\end{figure*}

@@

The cloud computing has been popular since 2006. Recently, there are growing interests in investigating the limits IaaS (Infrastructure as a Service) has faced. The IaaS is built on top of many servers and managed by a cloud operating system.  Cloud operating system accepts demands from users and allocate resources to them. While the interest in adopting this services is still growing, resource utilization of the CPU and memory is usually not full. The reason behind this under-utilization is that there is always segment issues for allocating CPU and memory. Furthermore, the servers underpinning the IaaS is not a good fit for raising trend of heterogeneous computing as adding various new computing devices into servers is not a trivial process. \par
    Hardware resource disaggregated architecture is proposed to address these limits. In this architecture, resources are no longer bound with a server. In other words, resources are collected in pools, such as processor pool, memory pool, and storage pool, and there is no boundary among each class of resource. In this setting, the resources can be distributed in the exact number of requested, the unallocated resources can be shutdown. \par
    While there is an operating system prototype, LegoOS, designed for the disaggregated architecture. We believe the main purpose of disaggregated architecture is to support cloud computing. Designing a hypervisor to support the IaaS functionality is therefore an important thing to be done. \par
    OmniVisor is a hypervisor in omni scale, for both aggregated and disaggregated architecture, and the goal is to address the limits of IaaS mentioned above. We foresee the practicable disaggregated architecture will not be complete in the recent years, so considering a hypervisor for aggregated architecture is also important. In supporting aggregated architecture, the difference between OmniVisor and previous hypervisors, such as Xen, KVM, and VMware’s products, are where does the computation happen. For previous hypervisors, all the computation is carried out within a server, i.e., an aggregated machine, while for OmniVisor, it makes computation happen across machines, i.e., cross-boundary computation. For example, OmniVisor is able to allocate processors from two servers and give them to a user. This resources separation is transparent to users, they still view it as a machine abstraction and use an operating system (guest operating system from the hypervisor’s point of view) to manage it. \par
    OmniVisor is a general solution to address limits of IaaS. It can be divided into two parts: support for aggregated and disaggregated architecture. It is certain that designing a hypervisor for disaggregated architecture will be an impactful work. As for the aggregated part, we should ask one question first: can the cross-boundary computation outperform current mechanism? For current design, even though the requested resources are more than the underlying machine can support, the hypervisor can still give out virtual resources by using time-sharing scheduler and storage to simulate processors and memory respectively. So we must first answer the benefits of cross-boundary computation in comparing with the previous design. \par
    This project will focus on the feasibility of OmniVisor for aggregated architecture. I will implement a prototype and evaluate the performance comparing with current hypervisors. Or at least, evaluate the performance requirement for designing the OmniVisor. For instance, since the cross-boundary computation requires communication through network, the network performance requirement for outperforming current hypervisors will be investigated. \par

@@

\emph{Poor hardware elastisity}

\emph{Poor failure handling}

@@

\begin{itemize}
\item \texttt{pdflatex latex}
\end{itemize}

@@

\subsection{Target Architecture}
For disaggregated architecture, all hardware computing resources such as processors, memory, and storage are separately located. Though processors and memory are separate, we assume there are still a few amount of memory stay with the processors. This assumption is reasonable given the network speed still not catches on the processor-memory performance requirement\cite{Network_requirement} and datacenter applications usually show great temporal data locality\cite{NN_load,MapReduce_load}. So in this paper, the memory components in our target architecture include local and remote memory as depicted in Fig \ref{fig:disaggregated architecture}.

We simulate the local and remote memory using the main memory of a machine as illustrated in Fig \ref{fig:local memory regions}. We manage the page table to separate the main memory as for simulating local and remote memory on the disaggregated architecture. Only the memory address to the local memory region is valid, accessing other address will cause a page fault. In the page fault handler, we will first check whether the page stays on the remote memory region and copy the data into local memory region if so. Otherwise, a storage access for reading the data into remote memory region, and then copy the data into local memory region. Local page eviction, page table entry invalidation, and TLB invalidation will also be executed to guarantee correctness. Moreover, we will execute a intentional wait in the page fault handler to simulate the latency for fetching a remote memory via network in the disaggregated architecture. The waiting time is a configurable variable in order to see in which level of the latency is acceptable for achieving comparable performance with monolithic architecture.

\subsection{Dedicated Local Memory}
In this paper, we explore the possibility of providing dedicated local memory to each virtual machine and compare it with the the shared local memory policy to see which is a better policy for disaggregated architecuture. See Fig \ref{fig:shared local memory} and Fig \ref{fig:dedicated local memory} for illustrations.

>>>>

\begin{figure*}[h!]
     \centering
     \captionsetup[subfigure]{position=b}
     \begin{subfigure}[b]{0.28\textwidth}
         \includegraphics[height=.2\textheight]{figures/monolithic.png}
         \caption{Monolithic architecture, hardware components are bounded within a machine.}
         \label{fig:monolithic architecture}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.68\textwidth}
         \includegraphics[height=.2\textheight]{figures/disaggregated.png}
         \caption{Disaggregated architecture, hardware components are separate: there are CPU-blade, memory-blade, storage-blade. We also assume there is local memory located with CPU}
         \label{fig:disaggregated architecture}
     \end{subfigure}
     \caption{Monolithic architecture and disaggregated architecture.}
     \label{fig:architecture}
\end{figure*}

@@

The cloud computing has been popular since 2006. Recently, there are growing interests in investigating the limits IaaS (Infrastructure as a Service) has faced. The IaaS is built on top of many servers and managed by a cloud operating system.  Cloud operating system accepts demands from users and allocate resources to them. While the interest in adopting this services is still growing, resource utilization of the CPU and memory is usually not full. The reason behind this under-utilization is that there is always segment issues for allocating CPU and memory. Furthermore, the servers underpinning the IaaS is not a good fit for raising trend of heterogeneous computing as adding various new computing devices into servers is not a trivial process.

Hardware resource disaggregated architecture is proposed to address these limits. In this architecture, resources are no longer bound with a server. In other words, resources are collected in pools, such as processor pool, memory pool, and storage pool, and there is no boundary among each class of resource. In this setting, the resources can be distributed in the exact number of requested, the unallocated resources can be shutdown. For the following article, we use monolithic architecture as the traditional server design and disaggregated architecture as our target architecture, as illustrated in Fig \ref{fig:architecture}.

While there is an operating system prototype, LegoOS\cite{LegoOS}, designed for the disaggregated architecture. We believe the main purpose of disaggregated architecture is to support cloud computing. Designing a hypervisor to support the IaaS functionality is therefore an important thing to be done.

For disaggregated architecture, all hardware computing resources such as processors, memory, and storage are separately located. Though processors and memory are separate, we assume there are still a few amount of memory stay with the processors. This assumption is reasonable given the network speed still not catches on the processor-memory performance requirement\cite{Network_requirement} and datacenter applications usually show great temporal data locality\cite{NN_load,MapReduce_load}. So in this paper, the memory components in our target architecture include local and remote memory as depicted in Fig \ref{fig:disaggregated architecture}.

This work is a pre-step before going into the design of hypervisor as there are some unclear questions about local memory on disaggregated architecture needed to be answered. We focus on local memory since it makes huge impact in the disaggregated architecture as the main memory stay remotely and has higher latency. First thing we explore is that, how many local memory is enough for running a virtual machine and to achieve comparable performance? Second, for running multiple virtual machines, should the local memory be dedicated or shared? This work uses a simulation approach to evaluate the performance of applications and try to answer the above questions.

We simulate the disaggregated architecutre by separating the main memory of a machine into local and remote memory regions. The processes are only allowed to access local memory region. page fault and page eviction happen if the data is not in local memory region. The data is first loaded into remote memory region and then is copied to local memory region before following access. A configurable delay is added in the page fault handler to simulate the latency in access remote memory via network.

To sum up, we believe building a hypervisor for disaggregated architecture is a more practical way to explore it. And in the virtualized environment, this paper explore what should be the right local memory design to achieve acceptable performance comparing with traditional monolithic architecture.

@@

In the traditional architecture, programs (for destop users) or virtual machine (as what IaaS provides) run on top of a monolithic machine where all of the hardware components are bounded within the machine.  While this setting has been working well for many years, current trends raise some distributeculties to adopt this model. Beside, there are also new techiniques to be enabling factors of disaggregated architecture.  The following sections discuss the limitation of monolithic servers and then descripe the primary concept of resource disaggregation.

\subsection{Monolithic Architecture Limitations}
In this subsection, we discuss about three limitations of monolithic architecture: rosource under-utilization, poor hardware elasticity, and poor failure handling.

First, in monolithic servers, all the hardware components are bounded within a physical machine. In the context of IaaS, for example, these machine boundaries are not a good fit for providing virtual machines to users. Think about a scenario that an user requests a virtual machine composed of many processors but only a few amount of memory. This request ends up using most of the processors but only a small portion of memory in a physical machine. It's then not possible for cloud OS to distribute other users' request to that physical machine as there are not many processors left. Therefore, the memory of the machine is under-utilized. This under-utilization issues could be generalized into any categoriy of hardware resources.

As for poor hardware elastisity, having boundary on hardware resources makes it extremely hard to add or remove hardware components to a machine. Increasing the server's computation power via adding more processors or memory is not easy to do. Furthermore, this limitation becomes more daunting as specific hardware accerlerators are viewed as a future trend these days\cite{Heterogeous_computing}, such as GPGPU\cite{Google_GPGPU}, TPU\cite{TPU}, FPGA\cite{Datacenter_FPGA}. While in disaggregated setting, any hardware component could be add or remove via attach or de-attach the HW-blade to other inter-connected hardware components.

Finally, the failure handling for monolithic architecure is that, even though there is a failure on only one hardware component, that failure usually causes entire system crushes. This is not an appealing outcome as people are expecting hardware component failures could be handled independently without breaking down the whole system. In the cloud provider, this failure model is even severe since the crush of entire machine will affect multiple users. In the disaggregated architecture, however, the faults of different components could be handled independently as there are loose-coupled.

@@

The concept of resource disaggregation is that instead of having hardware components bounded within a physical machine, different hardware components should be separate. We could have processors, memory, storages, and I/O devices sit in different place. These hardware components communicate with each other to provide functionalities that a computer should provide. From a user's persepctive, it just sees a machine abstraction carrying out all their computing requests.

Providing machine abstraction by separate hardware components face some technical challenges\cite{Rack_scale_challenges}. The first one is the efficiency of network communication. Performance degradation can be expected if the network speed is not fast enough. Another chanllenge is that traditional hardware components cannot work independently and might require some assistance from CPUs. Take DMA for example, though DMA is free of CPU's intervention during data copy, it still needs CPU to do some initialization. The cases are same in network and storage devices.

Fortunately, current techniques resolve some of the above challenges. We descipe some key enabling factors for resource disaggregation. First, faster network communication\cite{R2C2}, RDMA\cite{RDMA}, and InfiniteBand\cite{InfiniteBand} are introduced and becomes more and more popular recently. Furthermore, hardware devices with programmablity of standalone computing funcationality is also emerging\cite{PARDIS, PIM}.

@@

\subsection{Simulation Approach}
We simulate the local and remote memory using the main memory of a machine as illustrated in Fig \ref{fig:local memory regions}. We manage the page table to separate the main memory as for simulating local and remote memory on the disaggregated architecture. Only the memory address to the local memory region is valid, accessing other address will cause a page fault. In the page fault handler, we will first check whether the page stays on the remote memory region and copy the data into local memory region if so. Otherwise, a storage access for reading the data into remote memory region, and then copy the data into local memory region. Local page eviction, page table entry invalidation, and TLB invalidation will also be executed to guarantee correctness. Moreover, we will execute a intentional wait in the page fault handler to simulate the latency for fetching a remote memory via network in the disaggregated architecture. The waiting time is a configurable variable in order to see in which level of the latency is acceptable for achieving comparable performance with monolithic architecture. Also, the size of local memory region is changeable in order to do analysis on what local memory size is a better design for disaggregated architecture.

\subsection{Shared/Dedicated Local Memory}
In this paper, we explore the possibility of providing dedicated local memory to each virtual machine and compare it with the the shared local memory policy to see which is a better policy for disaggregated architecuture. See Fig \ref{fig:shared local memory} and Fig \ref{fig:dedicated local memory} for illustrations.
